### 2019-06-25 BGSM/Motivletter/Motivationletter.tex
BGSM stuff

\documentclass{letter}
\usepackage[spanish]{babel}
\usepackage[utf8]{inputenc}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}

\signature{Lázaro Alberto Larrauri Borroto}

\begin{document}
	\begin{letter}{}
		\opening{To Whom It May Concern:}
		\vspace{1cm}
		I am writing to submit my application for the Maria de Maeztu Master Internships Programme for
		2019. I am currently a student at the Master's degree in Advanced Mathematics and Mathematical
		Engineering offered by the Universitat Politécnica de Catalunya. I graduated from the University of
		Zaragoza in a Degree of Mathematics with various honours marks in subjects related to algebra, 
		discrete mathematics and geometry. I have participated in several math, physics and chemistry competitions 
		since secondary school and got good results in most of them.
		I came to know about the math masters programs offered here, in Barcelona, because of a friend 
		from those competitions, and I became interested. \par
	    My interest in research originated in secondary school, where I developed a strong liking and predisposition for
	    creative and intellectual activities. 
	    Eventually I decided to study math because I believe that it allows oneself to solve interesting, meaningful 
	    problems in creative ways. My main areas of interest are logic, discrete mathematics and algebra, 
	    as well as their applications in other fields, particularly computer science.  
		Recently, I got the opportunity to work on a master's thesis project under the supervision of Marc Noy Serrano,
		involving original research on logic, discrete mathematics and probability. I am very motivated by this project and I think
		it is a great chance for me to get more involved with research and to work hard doing things I enjoy.

		\closing{Yours Faithfully,}

	\end{letter}
\end{document}