### 2019-07-06 Presentacion Sparse Graphs/presentacio.tex

\documentclass[handout, 11pt]{beamer}

\usepackage[english] {babel}                 % Idioma (és important a l'hora de separar una paraula a final de línia)
%\input{myStyle.sty}

\usetheme{CambridgeUS}
\usepackage{color}
\usepackage{amssymb,mathtools, amsmath, amsfonts, amsthm}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\setbeamertemplate{itemize items}[circle]

\title[First Order Logic of Sparse Random Graphs]{First Order
Logic of Sparse Random Graphs}

\begin{document}
	\frame{\titlepage}

\begin{frame}{Preliminaries}
\begin{center}
	\large{First order language of graphs}
\end{center}
\begin{columns}
	\begin{column}{0.55\textwidth}
		\begin{itemize}
			\item Variables $x_1, \dots, x_n, \dots$
			\item Connectives $\wedge, \vee$, equality symbol $=$,
			and negation symbol $\neg$.
			\item Quantifiers $\forall, \exists$.
			\item A binary relation symbol $R$
		\end{itemize}
	\end{column}
		\begin{column}{0.45\textwidth}
		\begin{itemize}
			\item Vertices.
			\item ``And", ``or", ``equals", ``not".
			\item ``For all", ``there exists".
			\item Edges $x\sim y$. 
		\end{itemize}
	\end{column}
\end{columns}
~\\
\[ \forall x_1,x_2 \, R(x_1,x_2) \implies \exists x_3 \,
 \neg(x_3=x_1)\wedge \neg(x_3=x_1) \wedge R(x_1, x_3)\]
\end{frame}

\begin{frame}{Preliminaries}
	\Large{Binomial model of random graphs $G(n,p)$.} \\
	\large
	~\\
	Start with vertex set $[n]=\{1,\dots, n\}$ and
	each edge is added with probability $p$ independently. 
	\\~\\
	If $G=([n], E)$:
	\[ Pr(G)= p^{|E|}\cdot (1-p)^{\binom{n}{2}- |E|}. \]

\end{frame}

\begin{frame}{The problem}
For any F.O. sentence $\phi$ we can talk about
\[ Pr(G(n,p)\models \phi ), \]
And if $p=p(n)$ it also makes sense
\[ \Ln Pr(G(n,p(n))\models \phi) .\]

What do we know about these limits?

\end{frame}

\begin{frame}{The problem}	
For $p(n)=n^{-\alpha}$ and $\alpha\in [0,1]$ we have a nearly complete answer.
\begin{theorem}{(Fagin 19776)} If $0\leq p \leq 1$ is constant a Zero-One Law is satisfied
\end{theorem}
\begin{theorem}{(Spencer, Shelah 1988)} 
	\begin{itemize}
		\item If $\alpha \in (0,1)\setminus \mathbb{Q}$ then
		a Zero-One law holds.
		\item If $\alpha \in (0,1)\cup \mathbb{Q}$ then a Zero-One
		lay does not hold. Even more, there is a F.O. sentence
		$\phi$ such that
		\[ \Ln Pr(G(n,p(n))\models \phi) \]   
		does not exist.
	\end{itemize}
\end{theorem}

\end{frame}

\begin{frame}{The problem}
\begin{theorem}{(Lynch, 1992).} 
	If $p(n)=c/n$ for some positive real $c$
	the limit
	\[ \Ln Pr(G(n,p(n))\models \phi)\]
	always exists and is analytic in $c$.  
\end{theorem}
\end{frame}

\begin{frame}{The landscape of $G(n,c/n)$}
	We are interested in last theorem.\\~\\
	\begin{itemize}
		\item The number of cycles of 
		length $3,4\dots, r$
		are asymptotically distributed like independent Poisson 
		variables.
		\item Small cycles are a.a.s far away.
		\item Fixed vertices are a.a.s far away.
		\item The ball of a given radius centered in fixed
		vertex is a.a.s a tree. Any tree occurs with a positive 
		probability.
	\end{itemize}
\end{frame}

\begin{frame}{Outline of the proof}
	For any quantifier rank $k$ we show (constructively) that there exist
	some classes of graphs $C_1,.., C_m$ such that \\~\\
	\begin{itemize}
		\item[(1)] A.a.s any two random graphs in the same class
		have the same rank $k$ type.
		\item[(2)] A.a.s. a random graph belongs to some $C_i$.
		\item[(3)] $\Ln Pr(G\in C_i)$ exists for any $i$ and is
		analytic in $c$.
	\end{itemize}

\end{frame}
\begin{frame}{Outline of the proof}
	After this, for any F.O. sentence $\phi$ we obtain
	\\
	\begin{align*}
	 \Ln Pr(\phi) &= \sum_{i=1}^{m} \Ln Pr(\, \phi \wedge G\in C_i \,) + 
	  \Ln Pr(\, \phi \wedge G\notin C_1\cup \dots \cup C_l \,) \\
	 & \stackrel{(2)}{=} \sum_{i=1}^{m} \Ln Pr(\, \phi \, |\, G\in C_i \,)Pr( \, G\in C_i \,) \\
	 & \stackrel{(1)}{=} \sum_{j=1}^{l} \Ln Pr(\,G \in C_{i_j} \,)
	\end{align*}
	\\~\\
	And finally because of (3) this limit exists and is analytic in $c$.	
\end{frame}

\begin{frame}{Outline of the proof}

The classes $C_1,\dots, C_m$ are obtained as the equivalence classes
of a relation on graphs that \textbf{only} depends on the ``small" neighborhoods
of the small cycles. \\
\[ x \in  Core(G,r) \iff d(x,c)\leq r , \text{ for some cycle }c \text{ of length 
} \leq 2r+1 \]
~\\

\end{frame}

\begin{frame}{Outline of the proof}
\Large
	\begin{itemize}
		\item The probability of any given ``class" of core is given by
		a nested sequence of products of ``Poisson expressions¨, analytic on $c$.
		\\~\\
		\item To show that graphs with ``equivalent¨ cores share the same 
		rank $k$ type E.F games are used. 
	\end{itemize}

\end{frame}

\begin{frame}
	 \Large
	Can this be generalized to regular hypergraphs with edge size $a$ y
	edge probability $c/n^{a-1}$?\\~\\

	Yes it seems so. You can even consider arbitrary symmetry groups on the edges and
	multiple edge sets with different sizes.  
\end{frame}

\end{document}