### 2019-07-06 Presentacion/Presentación.tex

\documentclass[handout, 11pt]{beamer}

\usepackage[english] {babel}                 % Idioma (és important a l'hora de separar una paraula a final de línia)t
%\input{myStyle.sty}

\usetheme{CambridgeUS}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amssymb,mathtools, amsmath, amsfonts, amsthm}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\setbeamertemplate{itemize items}[circle]

\title[First Order Logic of Sparse Random Graphs]{First Order
Logic of Sparse Random Graphs}

\begin{document}
	\frame{\titlepage}
\begin{frame}
	\includegraphics[scale=0.5]{kmorphictrees}
\end{frame}

\end{document}