### 2019-07-07 Presentacion/Presentación.tex
Pictures

\documentclass[handout, 11pt]{beamer}

\usepackage[english] {babel}                 % Idioma (és important a l'hora de separar una paraula a final de línia)
%\input{myStyle.sty}

\usetheme{CambridgeUS}
\usepackage{color}
\usepackage{amssymb,mathtools, amsmath, amsfonts, amsthm}
\usepackage{graphicx}
\usepackage{float}
%\usepackage{hyperref}
%\usepackage{enumerate}
%\usepackage{enumitem}
\usepackage{chngcntr}
%\usepackage{cleveref}
\usepackage{pdfpages}
\setbeamertemplate{itemize items}[circle]

\newcommand{\cc}{\mathfrak{c}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Hc}{\mathcal{H}}
\newcommand{\Lan}{\mathcal{L}}
\newcommand{\Ln}{\lim\limits_{n\to \infty}}
\newcommand{\clist}{\mathfrak{c}_{1}, \cdots, \mathfrak{c}_m}
\newcommand{\morph}[1]{\stackrel{#1}{\simeq}}
\newcommand{\vlst}[2]{#1_1,\dots, #1_{#2}}
\newcommand{\gnp}{G(n,\beta_1/n^{a_1-1}, \dots,\beta_l/n^{a_l-1})}

\setbeamertemplate{itemize items}[circle]

\AtBeginSection[]
{
	\begin{frame}{Contents}
	\tableofcontents[currentsection]
	\end{frame}
}

\title[First Order Logic of Sparse Random Hyper-Graphs]{First Order
	Logic of Sparse Random Hyper-Graphs}

\author{L\'azaro Alberto Larrauri Borroto \\ 
	Supervisor: Marc Noy Serrano}

\date\today

\begin{document}
	\frame{\titlepage}
	\frame{\tableofcontents}

	\section{First Order Logic of Sparse Graphs}
	\begin{frame}{The first order language of graphs.}
		\begin{columns}
			\begin{column}{0.55\textwidth}
				\begin{itemize}
					\item Variables $x_1, \dots, x_n, \dots$
					\item Connectives $\wedge, \vee$, equality symbol $=$,
					and negation symbol $\neg$.
					\item Quantifiers $\forall, \exists$.
					\item A binary relation symbol $R$
				\end{itemize}
			\end{column}
			\begin{column}{0.45\textwidth}
				\begin{itemize}
					\item Vertices.
					\item ``And", ``or", ``equals", ``not".
					\item ``For all", ``there exists".
					\item Edges $x\sim y$. 
				\end{itemize}
			\end{column}
		\end{columns}
		~\\
		\[ \forall x_1,x_2 \, R(x_1,x_2) \implies \exists x_3 (\,
		\neg(x_3=x_1)\wedge \neg(x_3=x_2) \wedge R(x_1, x_3))\]
	\end{frame}

	\begin{frame}{The binomial model}
	The binomial model of random graphs $G(n,p)$ is a discrete probability space where 
	we assign to each graph $G=([n],E)$ the probability
	\[\mathrm{Pr}(G)= p^{|E|}\cdot (1-p)^{\binom{n}{2}- |E| }. \]
	\end{frame}

	\begin{frame}{Lynch's theorem}
	\begin{theorem}[Lynch, 1992]
		Let $\varphi$ be a sentence in the F.O. language of graphs. 
		Then the map $\mathrm{F}_\varphi: [0,\infty) \rightarrow \R$ given by 
		\[\mathrm{F}_\varphi(\beta)=\Ln \mathrm{Pr}(\, G(n,\beta/n)\models \varphi \,) \]
		is well defined and admits an analytic extension to $\CC$.
	\end{theorem}
	\end{frame}

	\begin{frame}{Overview of the proof}
	Some properties of $G(n,\beta/n)$:
	\\~\\
	\begin{itemize}
		\item The number of cycles of 
		length $3,4\dots, r$
		are asymptotically distributed like independent Poisson 
		variables.
		\item Small cycles are a.a.s far away.
		\item Fixed vertices are a.a.s far away.
		\item The ball of a given radius centered in fixed
		vertex is a.a.s a tree. Any tree occurs with a positive 
		probability.
	\end{itemize}

	\end{frame}

	\begin{frame}{Overview of the proof}
	For each fixed quantifier rank $k$:
	\\~\\
		\begin{itemize}
			\item[(1)]	It is given a finite classification of "small" uni-cycles.
			\item[(2)]  It is shown that the rank $k$ type of random graph $G$
			 in $G(n,\beta/n)$ a.a.s depends exclusively on the number of "small" uni-cycles
			 belonging to each class. 
			\item[(3)] The asymptotic distribution of those quantities is obtained. 
		\end{itemize}

	\end{frame}
	\section{General relational structures}
	\begin{frame}{Edge sets}
		\begin{definition} Fix
			\begin{itemize}
				\item natural numbers $n,a\in \N$, with $a\geq 2$
				\item a subgroup of the symmetric group $\Phi \leq S_a$
				\item and a subset of pairs $A\subset ([a]\times [a]) \setminus \Delta_a$.
			\end{itemize}
		The \textbf{total edge set} $\mathcal{H}_{(a,\Phi, A)}(n)$ of size $a$, 
		symmetry group $\Phi$ and
		restrictions $A$, on $n$ elements is the set
		\[ \mathcal{H}_{(a,\Phi, A)}(n)= ([n]^a/\Phi) \, \,
		\setminus R, \]
		where
		\[ R = \{\,  [x_1, \dots,x_a] \in [n]^a/\Phi  \, \, 
		| \, \, x_i=x_j \, \text{for some } (i,j)\in A \} \] 
		\end{definition}
	\end{frame}
	\begin{frame}{Graphs}
		\begin{definition}
			An (hyper)-graph $([n], H_1,\dots, H_c)$ with edge colors 
			$1,\dots, c$, sizes $a_1,\dots,a_c$, 
			symmetry groups $\Phi_1,\dots,\Phi_c$ and 
			restrictions $A_1,\dots,A_c$ consists of 
			\begin{itemize}
				\item The vertex set $[n]$ for some natural number $n$.
				\item For $i=1,\dots,c$, a ``colored" edge set $H_i\subseteq \mathcal{H}_{(a_i,\Phi_i,A_i)}(n)$ whose elements 
				have color $i$.
			\end{itemize}
		\end{definition}
	\end{frame}

	\begin{frame}{The first order language}
	Consider the first order purely relational language $\mathcal{L}$ whose signature consists of the relation symbols
	$R_1,\dots, R_c$ with arities $a_1,\dots,a_c$. \\~ \par

	A graph $G=([n],H_1,\dots, H_c)$ is a $\mathcal{L}$-structure in the following way:
	\begin{itemize}
		\item The universe of $G$ is its vertex set, $[n]$.
		\item For each $1\leq i \leq c$, 
		\[(x_1,\dots,x_{a_i})\in R_i^G \iff [x_1,\dots, x_{a_i}]\in H_i. \] 
	\end{itemize}
%		\begin{columns}
%			\begin{column}{0.55\textwidth}
%				\begin{itemize}
%					\item Variables $x_1, \dots, x_n, \dots$
%					\item Connectives $\wedge, \vee$, equality symbol $=$,
%					and negation symbol $\neg$.
%					\item Quantifiers $\forall, \exists$.
%					\item For each $1\leq i \leq c$ a relation symbol $R_i$ with arity $a_i$.
%				\end{itemize}
%			\end{column}
%			\begin{column}{0.45\textwidth}
%				\begin{itemize}
%					\item Vertices.
%					\item ``And", ``or", ``equals", ``not".
%					\item ``For all", ``there exists".
%					\item $i$-colored edges \\
%					$[x_1,\dots,x_{a_i}]$. 
%				\end{itemize}
%			\end{column}
%		\end{columns}
	\end{frame}

	\begin{frame}{The first order language}
		By definition, a graph $G=([n],H_1,\dots, H_n)$ satisfies, for each $1\leq i \leq c$:
		\vspace{0.5 em}
		\begin{itemize}
			\item Symmetry formulas:
			\[ S_g: =\big( R_i(x_1\dots,x_{a_i})\iff R_i(x_{g(1)}\dots,x_{g(a_i)})\big) ,\]
			where $g$ is an element from $\Phi_i$.
			\item Anti-reflexivity formulas:
			\[A_{i,(j,l)}:=\big(R_i(x_1\dots,x_{a_i})\implies 
			\neg(x_j= x_l)\big),\]
			where $(j,l)\in A_i$.
		\end{itemize}

	\end{frame}

	\begin{frame}{The random model}

	The random model $HG(n,p_1,\dots, p_c)$ is a discrete probability space where for each graph 
	$G=([n],H_1,\dots,H_c)$,
	\[ \mathrm{Pr}(G)=\prod_{i=1}^{c} p_i^{|H_i|}\cdot (1-p_i)^{|\mathcal{H}_{(a_i,\Phi_i,A_i)}(n)|-|H_i|} . \]
	We consider the case where for each $1\leq i \leq c$,  $p_i(n)=\beta_i/n^{a_i-1}$.
	\\~\\ 
	Let $\beta=(\beta_1,\dots, \beta_c)\in [0,\infty)^c$. We abbreviate
	\[HG(n,p(n,\beta)):= HG(n,\beta_1/n^{a_1-1},\dots, \beta_c/n^{a_c-1}). \]

	\end{frame}

	\begin{frame}{The theorem}
		We want to prove the following

		\begin{theorem} Let $\varphi$ be a first order sentence in $\mathcal{L}$.
			Then the map $\mathrm{F}:[0,\infty)^c\rightarrow \R$ given by
			\[ \mathrm{F}(\beta_1,\dots, \beta_c)= \Ln 
			\mathrm{Pr}(HG(n,p(n,\beta))\models \varphi )
			\]
			is well defined and admits an analytic extension to $\CC^c$. 
		\end{theorem}

	\end{frame}

	\begin{frame}{Distance and paths}
		Given any graph $G$, we define the following distance over its vertex-set:

			\begin{equation*}d(x,y)= \min_{\substack{H \leq G\\ 
					H \text{ connected }\\
					x,y\in V(H)}} \big(|V(H)| - 1 \big), \, \, \text{ or } \infty \text{ if }
				x,y \text{ are not connected.}
			\end{equation*}

		\begin{definition} 
			A path between two vertices $x,y$ in a graph $G$ is
			a connected subgraph $H\leq G$ containing both $x,y$ whose
			number of vertices is minimum. 
		\end{definition}

	\end{frame}
	\begin{frame}{Likelihood, trees, cycles and clusters} 
		\begin{definition}
			The likelihood $L(G)$ of a graph $G=(V,H_1,\dots,H_c)$ is
			the number
			\[ |V(G)| - \sum_{i=1}^c |H_i|(a_i-1).\]
		\end{definition}
	~\\
	\end{frame}

	\begin{frame}{Likelihood, trees, cycles and clusters} 
	\begin{itemize}
		\item A tree is a connected graph with likelihood $1$.
		\vspace*{0.5 em}
	\end{itemize}

	\begin{center}
		\includegraphics[width=0.7\linewidth]{Tree.pdf}
	\end{center}
	\end{frame}

	\begin{frame}{Likelihood, trees, cycles and clusters} 
		\begin{itemize}
			\item An unicycle is a connected graph with likelihood $0$.
			A cycle is a minimal unicycle.
		\end{itemize}
		\begin{center}
			\includegraphics[width=0.9\linewidth]{Cycles.pdf}
		\end{center}
	\end{frame}

	\begin{frame}{Likelihood, trees, cycles and clusters}
		\begin{itemize}
			\item A cluster is a graph $G$ with $L(G)\leq 0$ such 
			that $L(H)>L(G)$\\ for any subgraph $H\leq G$.
		\end{itemize}
	\begin{center}
		\includegraphics[width=0.4\linewidth]{Cluster.pdf}
	\end{center}
	\end{frame}

	\begin{frame}{The $k$-morphism relation over trees.}
		A rooted $(T,x)$ is a tree $T$ together with a distinguished vertex
		$x\in V(T)$.\par
		\[Tree(y, T) = (T[X], y),\]
		\[\text{where } \, X=\{ z\in V(T) \, | \, d(x,z)=d(x,y)+d(y,z)\}.\]~\\
		The root of an edge $e\in H(T)$ in a rooted tree 
		$(T,x)$ is the vertex $y\in e$ such that $d(x,y)=d(x,e)$.\\~\\
		The radius of an edge $e\in H(T)$ is 
		\[\max_{\substack{y\in e \\ y \text{ not the root of } e}} r(Tree(y,T)). \]

	\end{frame}

	\begin{frame}{The $k$-morphism relation over trees.}
		Fix $k\in \N$.\\
		We define the $k$-morphism relation over rooted trees of the same radius
		inductively as follows:

		\begin{itemize}
			\item If $r(T_1)=r(T_2)=0$ then $T_1\morph{k} T_2$.
		\end{itemize}
	\end{frame}

	\begin{frame}{The $k$-morphism relation over trees.}
		For $r>0$:
			\begin{itemize}
				\item  First we
				define the type of a en edge with radius less than $r$:

			\end{itemize}
		\begin{center}
			\includegraphics[width=0.8\linewidth]{Edgetype.pdf}
		\end{center}
	\end{frame}

	\begin{frame}{The $k$-morphism relation over trees.}
		\begin{itemize}
		\item If $r(T_1)=r(T_2)=r$ we say that $T_1\morph{K}T_2$ if
		for any edge $k$-type $E$ of radius less than $r$ either:
		\begin{itemize}
			\vspace{0.2 em}
			\item the number of initial edges in $T_1$ and $T_2$ of $k$-type $E$ 
			is the same, or
			\vspace{0.2 em}
			\item both $T_1$ and $T_2$ contain no less than $k+1$
			initial edges of $k$-type $E$. 
		\end{itemize}
		\begin{center}
			\includegraphics[width=0.7\linewidth]{Try.pdf}
		\end{center}

		\end{itemize}

	\end{frame}

\end{document}